\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usepackage{times}

\geometry{a4paper, margin=1in}
\title{Comprehensive Retail Analytics and Forecasting System}
\author{Arif Rasul}
\date{May 2025}

\begin{document}

\maketitle

\section{Executive Summary}
This project develops a Comprehensive Retail Analytics and Forecasting System using Python, SQLite, Streamlit, and Plotly to analyze retail sales, customer behavior, and inventory performance. The system manages a large dataset with 50,000 sales records across multiple tables, employs complex SQL queries for data extraction, and integrates machine learning for customer segmentation and sales forecasting. A multi-page Streamlit dashboard provides interactive visualizations and actionable insights, deployed on Streamlit Cloud for accessibility.

\section{Objectives}
The primary objectives are:
\begin{itemize}
    \item To build a scalable analytics system with SQL-driven data extraction.
    \item To implement advanced analytics, including K-means clustering and Prophet forecasting.
    \item To create a user-friendly, multi-page dashboard for sales, customer, and inventory insights.
    \item To demonstrate proficiency in Python, SQL, and data visualization for HR evaluation.
\end{itemize}

\section{Methodology}
\subsection{Data Source}
A synthetic dataset was generated, including:
\begin{itemize}
    \item Sales: 50,000 records (OrderID, CustomerID, ProductID, Date, Quantity, Price).
    \item Customers: 1,000 records (CustomerID, Age, Gender, Region).
    \item Products: 500 records (ProductID, Category, UnitCost).
    \item Inventory: 500 records (ProductID, StockLevel, RestockDate).
\end{itemize}
Data is stored in an SQLite database with indexes for performance.

\subsection{Technologies Used}
\begin{itemize}
    \item \textbf{Python}: Data processing (Pandas, NumPy), clustering (scikit-learn), forecasting (Prophet).
    \item \textbf{SQLite}: Database management with complex SQL queries.
    \item \textbf{Streamlit}: Multi-page interactive dashboard.
    \item \textbf{Plotly}: Dynamic visualizations.
    \item \textbf{GitHub}: Version control and deployment.
\end{itemize}

\subsection{Implementation Steps}
\begin{enumerate}
    \item \textbf{Database Setup}: Created an SQLite database with multiple tables and populated it with synthetic data.
    \item \textbf{Data Querying}: Developed SQL queries for joins, aggregations, and filtering across tables.
    \item \textbf{Data Processing}: Implemented Python functions for customer segmentation, inventory turnover, and forecasting.
    \item \textbf{Dashboard Development}: Built a multi-page Streamlit app with pages for overview, sales analysis, customer segmentation, inventory management, forecasting, and top customers.
    \item \textbf{Deployment}: Hosted the dashboard on Streamlit Cloud with a public URL.
\end{enumerate}

\section{Features}
\begin{itemize}
    \item \textbf{SQL Integration}: Over 10 complex queries for data extraction, including joins and aggregations.
    \item \textbf{KPIs}: Total sales, orders, average order value, customer retention, profit margin.
    \item \textbf{Visualizations}: Line charts, bar charts, pie charts, scatter plots, histograms, and forecast components.
    \item \textbf{Machine Learning}: K-means clustering for customer segmentation and Prophet for 60-day sales forecasting.
    \item \textbf{Inventory Management}: Turnover rates, low-stock alerts, and overstock detection.
    \item \textbf{Top Customers}: Analysis of high-value customers by revenue and order count.
    \item \textbf{Data Download}: Export filtered data as CSV.
\end{itemize}

\section{Outcomes}
\begin{itemize}
    \item Developed a robust system with over 600 lines of Python and SQL code.
    \item Demonstrated expertise in SQL, Python, machine learning, and visualization.
    \item Deployed a professional dashboard on Streamlit Cloud.
    \item Produced a comprehensive report for HR and portfolio inclusion.
\end{itemize}

\section{Challenges and Solutions}
\begin{itemize}
    \item \textbf{Challenge}: Managing large dataset with complex joins.
    \item \textbf{Solution}: Added indexes and optimized SQL queries.
    \item \textbf{Challenge}: Ensuring Streamlit Cloud compatibility.
    \item \textbf{Solution}: Specified exact dependency versions and Python 3.9 runtime.
    \item \textbf{Challenge}: Balancing dashboard complexity and usability.
    \item \textbf{Solution}: Organized into multi-page layout with intuitive navigation.
\end{itemize}

\section{Conclusion}
The Comprehensive Retail Analytics and Forecasting System showcases advanced skills in SQL, Python, machine learning, and web deployment. With extensive code and a polished dashboard, it is an ideal portfolio project to impress HR and demonstrate business impact.

\section{References}
\begin{itemize}
    \item Streamlit Documentation: \url{https://docs.streamlit.io}
    \item Prophet Documentation: \url{https://facebook.github.io/prophet}
    \item Plotly Documentation: \url{https://plotly.com/python}
    \item SQLite Documentation: \url{https://www.sqlite.org/docs.html}
    \item scikit-learn Documentation: \url{https://scikit-learn.org}
\end{itemize}

\end{document}